Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-09-08 Manuscript/manuscript.tex
Modified ternary complex binding plot for simulations -- now shows only simulations. Added a short discussion to it into SI in MS.
Started writing Binding stoichiometry section.

@@

>>>>

Binding stoichiometry of \ce{Ca2+} and POPC was 
thoroughly studied in the experimental work~\citenum{altenbach84}.
Several binding models were proposed and tested of which only one,
ternary complex binding model, provided a good fit of the experimental observations. 
In such a model, it is assumed that \ce{Ca2+} cations bind to a POPC membrane 
with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
Such a model also provides a good fit 
to our simulations with ECC-lipids 
(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
Simulations, however, allow us to directly evaluate the stoichiometry 
by calculating relative propensities of various \ce{Ca2+}:$n\times$POPC clusters
In Figure~\ref{fig:cacl_complexes} we see that ternary complex is indeed 
the most probable binding mode of calcium at 285~mM concentration.

Using our model, ECC-lipids, we fitted such a model to our simulations with a g
In such a model, there is a linear relationship between quantities 
$C_b$, mole fraction of bound \ce{Ca2+} per POPC, and $\sqrt(C_b/C_I)$, 
where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
The concentration $C_b$ was obtained from an extrapolation of linear relation 
between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
Such an extrapolation is valid as long as the mode of \ce{Ca2+} binding 
remains constant throughout the extrapolation range. 
The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
However, Boltzmann theory yields inaccurate results
for divalent cations like \ce{Ca2+}~\cite{Andelmann1995}. 
An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions. 
Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
our simulations likely lays in the fact that 
the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
gradually do not hold for higher concentrations of \ce{Ca2+}.

Binding probability:
\ce{Ca2+} complex with
1 POPC - 13\%
2 POPC - 18\%
3 POPC - 12\%

\begin{figure}[tbp]
  \centering
  \includegraphics[width=6.0cm]{../Fig/number_bound_lipids_p_cation_Ecc-lipids.eps} \\
  \caption{\label{fig:cacl_complexes}
      Histogram of relative probabilities of existence of complexes
      of \ce{Ca2+} and $n$ POPC lipids. 
  }
\end{figure}

@@

It was found in the original work~\cite{altenbach84} that 
a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
provides the best fit to experimental measurements of all considered models in that study. 
In such a model, there is a linear relationship between quantities 
$C_b$, mole fraction of bound \ce{Ca2+} per POPC, and $\sqrt{C_b/C_I}$, 
where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
The concentration $C_b$ was obtained from an extrapolation of linear relation 
between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
Such an extrapolation is valid as long as the mode of \ce{Ca2+} binding 
remains constant throughout the extrapolation range. 
The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
However, Boltzmann theory yields inaccurate results
for divalent cations like \ce{Ca2+}~\cite{Andelmann1995}. 
An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions. 
Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
our simulations likely lays in the fact that 
the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
gradually do not hold for higher concentrations of \ce{Ca2+}.

\begin{figure}[]
  \centering
  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
  \caption{\label{fig:cacl-bind}
    Ternary complex binding model of \ce{Ca2+} to a POPC membrane 
    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
    provides a good fit to experimental measurements~\cite{altenbach84}
    and it also provides a good fit to our simulation data. 
    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
    and, hence, the observed slope of the linear relationship is different.
    }
\end{figure}
